---
title: "2023_11_09_bulk_processing"
#format: html
editor: source
---

# Load Libraries

These libraries are required to perform this RNAseq analysis. Make sure you have install them before begin the analysis.

```{r}
suppressPackageStartupMessages({
library(tidyverse)
library(tidybulk)
library(janitor)
library(plotly)
library(ggrepel)
library(DESeq2)
library(GeneTonic)
library(pcaExplorer)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(topGO)
})
```


# Data Dictionary

Each sample was organized into its owned folder named using various variable separated by underscore ("\_"). For example, for"697_C\_CAR_Rep2_S5_L001":

-   697: cell line
-   C: treatment
-   CAR: condition
-   Rep2: replicate
-   S5: sequencing cell segment
-   L001: sequencing cell line

The code below manually listing samples' folder in a `vector` variable called `dir.names`

```{r}
dir.names = c(
"52-CD4_S37_L002",
"52-CD8_S38_L002",
"53-CD4_S39_L002",
"53-CD8_S40_L002",
"55-CD4_S41_L002",
"55-CD8_S42_L002",
"56-CD4_S43_L002",
"56-CD8_S44_L002",
"57-CD4-B_S45_L002",
"57-CD4-D_S46_L002",
"57-CD8-B_S47_L002",
"57-CD8-D_S48_L002",
"58-CD4-B_S49_L002",
"58-CD4-D_S50_L002",
"58-CD8-B_S51_L002",
"58-CD8-D_S52_L002",
"60-CD4-B_S53_L002",
"60-CD4-D_S54_L002",
"60-CD8-B_S55_L002",
"60-CD8-D_S56_L002",
"61-CD4_S57_L002",
"61-CD8_S58_L002",
"62-CD4_S59_L002",
"62-CD8_S60_L002",
"63-CD4_S61_L002",
"63-CD8_S62_L002",
"NK-1_S32_L002",
"NK-2B_S33_L002",
"NK-2D_S34_L002",
"NK-3_S35_L002",
"NK-4_S36_L002"
)
```